\section{MODULE VII.}
\subsection{Exploration of the solar system.} 
The early missions to the Moon, Mars, Venus and Mercury. The Apollo project. Missions to the outer planets. Robotic exploration of Mars. New Horizons, Rosetta and Dawn missions. Gravity assisted interplanetary navigation.

Complementary information: embedded in the corresponding chapters.

At the end of this module you should be able to:
\begin{itemize}
\item  Appreciate the value of both robotic and human exploration of the solar system towards the understanding of mainly the origin and development of life
\item Understand the need for international collaboration
\item Appreciate the variety of missions along more than 50 years
\item Identify the major challenges to the human exploration of the solar system
\item Understand the principle behind the Lagrange points and their practical application
\item Understand the mechanics of gravity assisted navigation including a few important examples of its use, specially the Rosetta mission
\end{itemize}

ADDITIONAL REFERENCE MATERIAL I:
\begin{itemize}
\item Universe, Freedman. Chapters 14 and 15
\item \url{http://en.wikipedia.org/wiki/MAVEN}
\item \url{http://www.nasa.gov/mission_pages/dawn/main/#.VRrpJY6LXkY}
\item \url{https://solarsystem.nasa.gov/missions/profile.cfm?MCode=Hayabusa2}
\item \url{http://www.esa.int/Our_Activities/Space_Science/Rosetta}
\item \url{https://www.nasa.gov/mission_pages/deepimpact/main/} 
\item \url{http://global.jaxa.jp/projects/sat/muses_c/}
\item \url{http://www.nasaspaceflight.com/}
\end{itemize}

ADDITIONAL REFERENCE MATERIAL II:

There are hundreds of excellent references on the internet about space exploration. Wikipedia is very good as basic source for any space mission. Specifically Apollo, Rosetta, Pioneer, Voyager, Mars rovers, soviet Luna project, DAWN and New Horizons.

Note how in general terms, the science aims at detecting presence of water and environments suitable for life (Mars, moons, asteroids and comets).

General references:
\begin{itemize}
\item \url{http://cosmicdiary.org/fmarchis/files/2014/05/50-years-of-exploration.jpg}
\item \url{http://harold-clenet.com/site/wp-content/uploads/2013/02/Mars_Mission_Update.jpg}
Pioneer plaque and Voyager disc decoded: \url{http://en.wikipedia.org/wiki/Pioneer_plaque} 
\item \url{https://en.wikipedia.org/wiki/Voyager_Golden_Record}
\end{itemize}